---
title: "Introducing synopR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing synopR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
**2026-03-04**
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


**synopR** has been designed to easily transform SYNOP messages into tidy data frames ready for subsequent analysis.

## Workflow with Ogimet
Decoding with **synopR** is straightforward. Let's start with SYNOP messages retrieved from [Ogimet](https://www.ogimet.com/cgi-bin/getsynop?block=87736&begin=202602010300&end=202602012300) for the Rio Colorado station, Argentina (WMO identifier: 87736). After downloading and saving the data as a .txt file, we can read it using readLines().

```{r carga-datos}
# Temporary file
synop_file <- tempfile(fileext = ".txt")
writeLines(c("87736,2026,02,01,03,00,AAXX 01034 87736 NIL=",
"87736,2026,02,01,06,00,AAXX 01064 87736 NIL=",
"87736,2026,02,01,09,00,AAXX 01094 87736 NIL=",
"87736,2026,02,01,12,00,AAXX 01123 87736 12965 31808 10240 20210 39992 40082 5//// 60104 82075 333 10282 20216 56055 82360=",
"87736,2026,02,01,15,00,AAXX 01154 87736 NIL=",
"87736,2026,02,01,18,00,AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818=",
"87736,2026,02,01,21,00,AAXX 01214 87736 NIL="), synop_file)

# Read the file
data_input <- data.frame(synops = readLines(synop_file))

print(data_input)
```

Take a look at the data. We can observe two specific characteristics. First, there are NIL messages, which contain no meteorological data. Second, these are not "pure" SYNOP strings; they include a prefix added by Ogimet that specifies the station ID (87736) along with the date and time of the observation.

However, this is not an issue, as we can use the `parse_ogimet()` function. This tool is specifically designed to strip away these headers and extract the raw SYNOP message for processing.

```{r parse_ogimet}
library(synopR)

# Note that `parse_ogimet(data_input)` is incorrect
data_from_ogimet <- parse_ogimet(data_input$synops) 

print(data_from_ogimet)
```

Before decoding, it is good practice to check the SYNOP messages for non-standard structures. Occasionally, errors occur where data groups are not separated by spaces or contain only 4 digits instead of the required 5. The `check_synop()` function is designed to handle these cases, ensuring the integrity of the data before processing.

```{r check synops}

check_synop(data_from_ogimet$Raw_synop)

check_synop(data_from_ogimet)

```

The `check_synop()` function accepts either a character vector or a specific data frame column containing SYNOP strings. A data frame with multiple columns —where the SYNOP column is not explicitly specified— will be accepted **if and only if that data frame is the direct output of** `parse_ogimet()`.

```{r check synops2}

my_df <- data.frame(syn = c("AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818=",
                            "AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818="),
                    second_column = c(5,7))

check_synop(my_df)

check_synop(my_df$syn)

```

So far, our messages have a correct structure (even the NIL ones). Now, let’s see what happens when they don't.

```{r check synops3}

check_synop(c("AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818=",
              "AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 6000182100 333 56600 82818=",
              "AAXX 01183 87736 12465 20000 10326 2021 39974 40064 5//// 60001 82100 333 56600 82818=",
              "AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818",
              "Not a synop message="))

```

`check_synop()` returns a tibble where the first column indicates whether each SYNOP is valid (TRUE) or not (FALSE), and the second column describes the specific error found.
In our example:

* The first SYNOP is correct.
* In the second, there is a missing space between groups 6 and 8 in Section 1.
* In the third, group 2 of Section 3 contains only 4 digits.
* The fourth message is missing the "=" terminator (remember that SYNOP messages must always start with "AAXX" and end with "=").
* The fifth is simply not a SYNOP string at all.

Now, we are ready to extract the information contained in the messages. The `show_synop_data()` function is where the real work begins.

```{r extraccion datos}

my_data <- show_synop_data(data_from_ogimet, wmo_identifier = '87736')

knitr::kable(t(my_data))
```

The `wmo_identifier` argument might seem like an extra step, but it offers a significant advantage: it allows for automatic filtering in case the data contains messages from other stations.

While the following example uses a vector with only two messages for simplicity, if you are working with thousands of SYNOP strings from multiple stations, this built-in filtering becomes extremely convenient.


```{r ventaja wmo_identifier}

# Messages from 87736 and 87016
mixed_synop <- c("AAXX 01183 87736 12465 20000 10326 20215 39974 40064 5//// 60001 82100 333 56600 82818=",
                 "AAXX 04033 87016 41460 83208 10200 20194 39712 40114 50003 70292 888// 333 56699 82810 88615="
                 )

colorado_data <- show_synop_data(mixed_synop, wmo_identifier = '87736')
print(colorado_data)
```

## Standard workflow
All `show_synop_data()` requires is a character vector or a data frame column where each element is a SYNOP string.
 
```{r extraccion datos estandar}

data_input_vector <- c("AAXX 04003 87736 32965 00000 10204 20106 39982 40074 5//// 333 10266 20158 555 64169 65090 =",
                       "AAXX 01094 87736 NIL=",
                       "AAXX 03183 87736 32965 12708 10254 20052 30005 40098 5//// 80005 333 56000 81270 =")

my_data <- show_synop_data(data_input_vector, wmo_identifier = '87736')

print(my_data)

```

Group 555 (reserved for national distribution) is currently ignored, as its content varies by country. However, future versions of **synopR** may include functions to extract data from this section based on user requirements.

## Limitations
**synopR** don't accept sections 222 y 444. Please visit the official [GitHub](https://github.com/ezequiel1593/synopR) repository for more information regarding limitations and assumptions.

